---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(tidyverse)
```

# vermeulen

<!-- badges: start -->

<!-- badges: end -->

`{vermeulen}` provides the Biomarker data set by [Vermeulen et al.
(2009)](https://doi.org/10.1016/S1470-2045(09)70154-8) in tidy format.

This data set is for a real-time quantitative PCR experiment that comprises:

-   The raw fluorescence data of 24,576 amplification curves.
-   64 targets: 59 genes of interest and 5 reference genes.
-   366 neuroblastoma cDNA samples and 18 dilution series samples.

## Installation

Install `{vermeulen}` from CRAN:

``` r
# Install from CRAN
install.packages("vermeulen")
```

You can instead install the development version of `{vermeulen}` from GitHub:

``` r
# install.packages("remotes")
remotes::install_github("ramiromagno/vermeulen")
```

## Usage

Because of CRAN size limits the data is not provided at installation time. The
data can be retrieved from this GitHub repository after installation with the
function `ds_biomarker()`.

```{r}
library(vermeulen)
library(tidyverse)

# Takes a few seconds (downloading from GitHub...)
biomarker <- as_tibble(ds_biomarker())
biomarker
```

Types of samples:

```{r}
count(
  distinct(biomarker, plate, well, sample_type, copies, dilution),
  sample_type,
  copies,
  dilution
)
```

Alternatively, if you want only a selection of variables you may use the 
following helpers:

- `amplification_curves()`
- `reactions()`
- `targets()`
- `samples()`

## Code of Conduct

Please note that the `{vermeulen}` project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

## References

1.  Vermeulen et al.. *Predicting outcomes for children with neuroblastoma using a multigene-expression signature: a retrospective SIOPEN/COG/GPOH study*. The Lancet Oncology 10, 663--671 (2009). doi: [10.1016/S1470-2045(09)70154-8](https://doi.org/10.1016/S1470-2045(09)70154-8).

2.  Ruijter et al.. *Evaluation of qPCR curve analysis methods for reliable biomarker discovery: Bias, resolution, precision, and implications*. Methods 59 32--46 (2013). doi: [10.1016/j.ymeth.2012.08.011](https://doi.org/10.1016/j.ymeth.2012.08.011).
